\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{xcolor}
\usepackage{multirow}
% Preamble (one-time)
\usepackage[most]{tcolorbox}
\tcbuselibrary{listings,breakable,skins}

% A full-width (double-column) prompt box with nice corners + background
\newtcblisting{FullWidthPromptBox}[2][]{%
  enhanced,
  breakable,
  float*=t,            % uses figure* to span both columns in twocolumn
  width=\textwidth,
  colback=gray!6,      % background
  colframe=black,
  colbacktitle=gray!18,
  coltitle=black,
  fonttitle=\bfseries,
  boxrule=0.9pt,
  arc=3mm,             % rounded corners
  outer arc=3mm,
  drop shadow,
  left=3mm,right=3mm,top=2mm,bottom=2mm,
  listing engine=listings,
  listing only,
  title={#2},
  listing options={
    basicstyle=\ttfamily\footnotesize,
    breaklines=true,
    breakatwhitespace=false,
    columns=fullflexible,
    keepspaces=true,
    showstringspaces=false
    % Uncomment if you want line numbers:
    % ,numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=6pt
  },
  #1
}


\bibliographystyle{IEEEtran}


\newcommand{\ourmethod}[1]{\textcolor{black}{HALO}}
\newcommand{\rutav}[1]{\textcolor{blue}{RS: #1}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\yuke}[1]{\textcolor{purple}{#1}}
\newcommand{\roberto}[1]{\textcolor{green}{#1}}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Memory Retrieval in Visuomotor Policies for Long-Horizon Robot Control}
% \title{Learning to Retrieve Past Information for Diverse, Long-Horizon Robot Manipulation Tasks}
% \title{Learning to Retrieve from History for Robot Manipulation via Knowledge Distillation}
%\title{Retrieval-Augmented Visuomotor Policies for Diverse, Long-Horizon Manipulation Tasks}
% \title{Retrieval-Augmented Visuomotor Policies for Imitation Learning under Partial Observability}
% \title{Retrieval-Augmented Visuomotor Policies for Long-Horizon Imitation Learning under Partial Observability}
% \title{Retrieval-Augmented Policies for Long-Horizon Imitation Learning under Partial Observability}
% \title{Memory-Augmented Visuomotor Policies using Learned Retrieval for Long-Horizon Robot Control}

% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
% should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}


% Author Names Omitted for Anonymous Review. Paper-ID: 348
\author{
Rutav Shah$^{1}$, 
Yisu Li$^{*}$, 
Femi Bello$^{*}$, 
Yuke Zhu, 
Roberto Mart{\'i}n-Mart{\'i}n \\
\vspace{-8pt}
\thanks{$^1$ Corresponding author: rutavms@utexas.edu, $^{*}$ Equal contribution} \\
The University of Texas at Austin
\vspace{-5pt}
}


\maketitle

\begin{abstract}
General-purpose robots operating in partially observable environments, such as homes, require memory to support autonomy.
They must recall diverse information from the past, such as where objects were placed, which tasks a human partner has completed, and when an appliance was turned on, to accomplish a wide range of tasks.
Achieving this versatility requires a memory retrieval mechanism that generalizes well across tasks.
However, hand-designed or heuristic-based methods rely on task-specific assumptions that may not transfer to different settings.
Transformer architectures that use attention over long contexts for memory retrieval provide a promising alternative, as they learn retrieval from data without task-specific assumptions.
However, directly incorporating long-context transformer architecture into imitation learning from offline data introduces two key challenges: (1) the policy may learn spurious correlations between the information from the past and predicted actions, and (2) errors accumulate over time in the memory due to prediction inaccuracies and their compounding interactions with the environment, leading to model drift and cascading failures in long-horizon control.
To address both challenges, we introduce \ourmethod{}, a visuomotor policy with an attention-based memory retrieval mechanism for long-horizon control.
To suppress spurious correlations, \ourmethod{} leverages vision-language model (VLM) priors to steer retrieval toward task-relevant information. Concretely, it generates task-relevant, memory-dependent question--answer pairs from demonstration trajectories and trains the policy jointly with a video question-answering objective, transferring VLM priors to the visuomotor policy.
Second, to reduce the impact of accumulated errors in memory during closed-loop control, \ourmethod{} uses sparse attention that restricts retrieval to only the most relevant parts of the history.
Together, these components enable more reliable long-horizon control by guiding the policy to retrieve task-relevant information from up to eight minutes of past experience.
Project website: \url{https://robin-lab.cs.utexas.edu/HALO}
% General-purpose robots operating in household environments \yuke{make the sentence more logical: it is because they must operate in partially observable environments for long-term autonomy -- thus they need memory} must recall different types of information from the past, such as where objects were placed, which subtasks have already been handled by their human partners, and when the appliance was turned on.
% This capability requires an effective memory retrieval mechanism, where hand-designed or heuristic-based retrieval methods often fail to generalize \yuke{break this down into two sentences as you are conveying two ideas: 1) memory retrieval is needed, and 2) hand-designed and heuristic methods won't work}.
% To mitigate spurious correlations, \ourmethod{} leverages knowledge from vision-language models (VLMs) to \yuke{[this is vague -- how exactly did you do it] induce task-specific priors for memory retrieval.} 
% To reduce \yuke{[jargon - hard to understand] accumulated model drift}, \ourmethod{} employs sparse attention mechanisms that restrict retrieval to the most relevant portions of the history.
% Together, these components enable robust long-horizon control, allowing robots to retrieve \yuke{[should not just be diverse (?), is relevance more important?] diverse} forms of information from up to two minutes of past experience.
\end{abstract}

% \begin{IEEEkeywords}
% memory retrieval, visuomotor policies, imitation learning.
% \end{IEEEkeywords}

\input{01_introduction}
\input{02_related_works}
\input{03_method}
\input{04_experiments}
\input{05_conclusion}
\input{06_acknowledgements}
% \clearpage
\bibliography{references/neuroscience, references/ai, references/appendix}

\clearpage
\input{Appendix}
% \pagebreak
% \pagebreak
% \pagebreak
%\bibliography{references/appendix}
% \bibliography{references/appendix, references/neuroscience, references/ai}
\end{document}
